---
title: "Topic ideas"
author: "LAM-duh: Xuliang Deng, Leah Okamura, Megan Richards"
date: "Oct 11th, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```


```{r load-packages}
library(tidyverse) 
# load other packages as needed
``` 

```{r load-data}
# load all data sets 
pokemon <- read_csv('data/Pokemon.csv')
hotels <- read_csv("data/hotels_reduced.csv")
schools <- read_csv("data/US_School_Racial_Segregation.csv")
```

# Data Set 1

## Introduction and Data

The dataset we are using is from https://github.com/ayoubimaya/pokemon and is sourced from https://www.pokemon.com/us/pokedex/, which is the official Pokemon open database. In this dataset, every observation is a specfic pokemon from the pokedex. The general characteristics being measured in the data are the name, type, and characteristics of each Pokemon. More specifically, some variables that we are interested in are Attack, HP, Defense, Speed, and the Pokemon type.

## Research questions 

We would like to see if there are any differences in the means of attack, defense, HP, based on each type using an ANOVA test. We are the interested in seeing how these means might differ to each other through hypothesis testing. We are also interested in analyzing how variables such as attack, defense, and speed may affect the total points assigned to a specific pokemon.

# Data Set 2

## Introduction and Data

This dataset is found at  https://datacatalog.urban.org/dataset/index-school-contribution-racial-segregation-us-school-districts, and is available through the Urban Institute. It contains information on the contribution of individual schools to the racial and ethnic segregation of US school districts using the SCI. The Segregation Contribution Index (SCI) measures the share of school district segregation attributable to a given school. The SCI is computed using public data on school enrollment by race and ethnicity from the US Department of Education’s Common Core of Data and the Private School Survey for the 2017–18 school year. It contains the following variables of interest:

School Level, Longitude of school location, Latitutde of school location, Charter School Indicator, Private School Indicator, Magnet School Indicator, Traditional Public School (TPS) indicator, Total school enrollment, Black and Hispanic school enrollment, Size of School, Neighborhood Radius (miles), Number of neighborhood schools serving same grades, Total enrollment in neighborhood, Black and Hispanic school enrollment in neighborhood, Urban school indicator (NCES), Suburban school indicator (NCES), Town / Rural school indicator (NCES), Segregation Contribution Index (SCI). 

## Research questions 

We are interested in investigating how well the following factors predict the the SCI (Segregation
Contribution Index), which is a measure of the share of a school district's segregation that 
is attributable to the school: 
- location (longitude and latitute), 
- school status (charter, private, magnet)
- school population 
- demographic make-up 
- the neighborhood radius
- the number of district schools with the same grading scores
- the neighborhood population
- the demographic make-up of the school district neighborhood 
- the region of the school (urban, suburban, rural)

We are also interested in comparing the mean SCI for the grouped variables (such as 
urban/suburban/rural), and conducting ANOVA tests to evaluate group variance. 

# Data Set 3

## Introduction and Data

The source of the dataset is Tiny Tuesday, https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md.
This data set comes from an open hotel booking demand dataset from Antonio, Almeida and Nunes, 2019. It is sourced from this study https://www.sciencedirect.com/science/article/pii/S2352340918315191#f0010. Due to the dataset being over 100,000 observations, we have limited the observations to be only hotels from the US. The general characteristics being measured in the data are the different aspects of booking and staying at a hotel. For example, out of the 32 variables, some of the ones we find great interest in are hotel type, reserved room type, assigned room type, company, meal, number of adults/children/babies, the average daily rate or daily cost, and the reservation status. 

## Research questions 

A research question we are interested in is how do factors such as type of hotel and type of guest affect the average daily rate for a hotel. We would also be interested in seeing how stays in the weekend or the weekday may affect the average daily rate for a hotel, and if they differ between the two hotel types, City and Resort hotels.

# Glimpse of data sets

## Data set 1

```{r glimpse-1}
glimpse(pokemon)
```

## Data set 2

```{r glimpse-2}
glimpse(schools)
```

## Data set 3

```{r glimpse-3}
glimpse(hotels)
```

