---
title: "Topic ideas"
author: "LAM-duh: Xuliang Deng, Leah Okamura, Megan Richards"
date: "Update date"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```


```{r load-packages}
library(tidyverse) 
# load other packages as needed
``` 

```{r load-data}
# load all data sets 
salary <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv')
hotels <- read_csv("data/hotels_reduced.csv")

```

# Data Set 1 - XULIANG

## Introduction and Data

(Introduce and discuss data here)
"Ask a Manager Salary"
https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-05-18/readme.md
The source of the dataset is Tiny Tuesday from the link above. The data is collected based on the survey done online on this website, https://www.askamanager.org/2021/04/how-much-money-do-you-make-4.html. The survey was about salary of managers in the US. The general characteristics being measured in the data are most of the managers who answered this survey are white female and mostly college graduates and shows relationships between certain factors and managers' income. 
Since some data includes N/A or characters, we need to clean the data at the beginning for sure.
## Research questions 

(Discuss research questions here)
We would like to see how manager salary is influenced or differs depending on other factors such as places managers live in, age, years in the job, highest level of education, gender, industry, and race. Since the data is not clean, for instance for race, we can only compare white to anything else. Race can be "African American or Asian American" and it might be hard to filter data.

# Data Set 2 - MEGAN

## Introduction and Data

(Introduce and discuss data here)

## Research questions 

(Discuss research questions here)



# Data Set 3 - LEAH
## Introduction and Data

The source of the dataset is Tiny Tuesday, https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md.
This data set comes from an open hotel booking demand dataset from Antonio, Almeida and Nunes, 2019. It is sourced from this study https://www.sciencedirect.com/science/article/pii/S2352340918315191#f0010. Due to the dataset being over 100,000 observations, we have limited the observations to be only hotels from the US. The general characteristics being measured in the data are the different aspects of booking and staying at a hotel. For example, out of the 32 variables, some of the ones we find great interest in are hotel type, reserved room type, assigned room type, company, meal, number of adults/children/babies, the average daily rate or daily cost, and the reservation status. 

## Research questions 

A research question we are interested in is how do factors such as type of hotel and type of guest affect the average daily rate for a hotel. We would also be interested in seeing how stays in the weekend or the weekday may affect the average daily rate for a hotel, and if they differ between the two hotel types, City and Resort hotels.


# Glimpse of data sets

## Data set 1 - XULIANG

```{r glimpse-1}
glimpse(salary)
```


## Data set 2 - MEGAN

```{r glimpse-2}

```


## Data set 3 - LEAH

```{r glimpse-3}
glimpse(hotels)
```

