---
title: "Topic ideas"
author: "LAM-duh: Xuliang Deng, Leah Okamura, Megan Richards"
date: "Update date"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)
```


```{r load-packages}
library(tidyverse) 
# load other packages as needed
``` 

```{r load-data}
# load all data sets 
survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv')

```

# Data Set 1 - XULIANG

## Introduction and Data

(Introduce and discuss data here)
"Ask a Manager Salary"
https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-05-18/readme.md

This data is from the salary survey to managers. The information included in the data can be seen from the github link above.
Since some data includes N/A or characters, we need to clean the data at the beginning for sure.
## Research questions 

(Discuss research questions here)
I would like to see how manager salary is influenced or differs depending on other factors such as places managers live in, age, years in the job, highest level of education, gender, industry, and race. Since the data is not clean, for instance for race, we can only compare white to anything else. Race can be "African American or Asian American" and it might be hard to filter data.

# Data Set 2 - MEGAN

## Introduction and Data

(Introduce and discuss data here)

## Research questions 

(Discuss research questions here)



# Data Set 3 - LEAH
## Introduction and Data

(Introduce and discuss data here)

## Research questions 

(Discuss research questions here)



# Glimpse of data sets

## Data set 1 - XULIANG

```{r glimpse-1}
survey
```


## Data set 2 - MEGAN

```{r glimpse-2}

```


## Data set 3 - LEAH

```{r glimpse-3}

```

