---
title: "Predicting the Cost of Hotel Booking in the United States"
author: "LAM-duh: Xuliang Deng, Leah Okamura, Megan Richards"
date: "Dec 6th, 2021"
output: 
  pdf_document: 
    toc: true
    toc_depth: 1
    number_sections: true

bibliography: references.bib
link-citations: true
---


```{r ggplot2-options, include = FALSE, echo = FALSE, message=FALSE}
# set consistent theme
#theme_set(ggplot2::theme_minimal())
color_palette <- list(gray = "#999999", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      purple = "#800080", 
                      orange = "#ff8c00",
                      cyan = "#008B8B"
                      )
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages,include=FALSE, echo = FALSE,message=FALSE}
library(tidyverse)
library(patchwork)
library(knitr)
library(tidymodels)
library(ggfortify)
library(ggplot2)
#install.packages("Rpdb")
``` 

```{r load-data,include=FALSE, echo = FALSE, message=FALSE}
hotels <- read_csv("data/hotels_reduced.csv")
```

# Introduction/Data 

## Introduction 
Hotels are a critical component of the travel sector in the U.S., with an estimated 5.3 million guest rooms, and supporting 1 in 25 American jobs on average prior to 2020 based on an [American Hotel and Lodging Association Report](https://www.ahla.com/sites/default/files/2021_state_of_the_industry_0.pdf). [A 2018 report](https://www.trekksoft.com/en/blog/65-travel-tourism-statistics-for-2019) showed that 
approximately two thirds of Americans book their hotels directly through hotel websites, and [a 2019 analysis](https://globalnews.booking.com/bookingcom-reveals-key-findings-from-its-2019-sustainable-travel-report/) showed large discrepancies in consumer travel cost based on how consumers chose to prioritize booking lodging compared to activities.

We are interested in analyzing U.S. hotel bookings, with the goal of providing insight to consumers on the factors that affect their average daily hotel cost. We believe there will be several significant points of relevance for understanding these relationships: understanding predictors of room cost could be used to help travelers to plan financially for future travel, or to potentially reduce cost. In this report, we are looking to use a variety of chosen models to understand the contributing factors to the average daily rate of a hotel room, as well as identify the strongest predictors.

To better understand the contributing factors to the daily rate of hotel rooms in the U.S., we build models to:

1) Evaluate the efficacy of this model to predict daily hotel room rates
2) Evaluate the statistical significance of each predictor, identifying the strongest contributing factors. 

We find a model that is a reasonable predictor of daily hotel room rates in the U.S., 
and find that purchase of a meal plan, booking in August or October, and having previous
bookings that were not cancelled were the most correlated with larger increases in daily hotel rate, 
and that being a repeated guest, booking in a Resort hotel instead of a City hotel, 
and booking in December were most correlated with decreases in daily hotel rate.
We were also able to show that our model generalized to new
data, which means that it could reasonably be used to predict hotel rates in the U.S., 
with an expected error of about $33. 

## Data 

Our data set is hosted on [Tiny Tuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md), was originally collected through an global open hotel booking demand dataset [@ANTONIO201941], and was originally analyzed in a [2019 study](https://www.sciencedirect.com/science/article/pii/S2352340918315191#f0010). Due to the dataset being over 100,000 observations, we have limited the observations to be only hotels from the US. 

The dataset contains characteristics of hotel bookings and hotel stays. Each observation represents one booking/stay at a specific hotel in America, containing information about the hotel itself, the booking details, and the 
occupants. The full data set dictionary can be found in the repository ReadMe. For this analysis, we evaluate
the following predictors: hotel type, reserved room type, assigned room type, company, meal, number of adults/children/babies, the average daily rate or daily cost, and the reservation status. Our methodologies section includes information on predictor selection. 

## Exploratory Data Analysis 

We start our analysis by exploring the data. Since we are interested in predicting 
the average daily rate, we first plot its distribution and calculate its summary
statistics. 


```{r response-plot, echo = FALSE, message=FALSE}
ggplot(data = hotels, aes(x = adr)) +
  geom_histogram(fill = 2) + 
  labs(x = "Price in US Dollars",
       y = "Frequency", 
       title = "Distribution of Average Daily Rate (Cost) of Hotel Bookings",
       subtitle = "Collected from Hotels in the U.S. from 2015-2017") 

```

```{r summary stats, echo = FALSE, message=FALSE}
hotels %>%
    summarise(mean = mean(adr), 
              median = median(adr), 
              sd = sd(adr),
              min = min(adr),
              max = max(adr),
              iqr = IQR(adr))%>%
kable(digits = 3)
```

The response variable, adr, has a somewhat skewed right, bimodal distribution. The average or mean average daily rate is \$122.992 and the median is $115. Because the distribution is skewed, the median is most likely the best indicator for the center. The standard deviation is \$51.617 and the data ranges from \$0 to \$328.33 with an interquartile range of \$61.99.

```{r, echo = FALSE, message=FALSE}
ggplot(data = hotels, aes(x = hotel, fill = hotel)) +
geom_bar() +
labs(title = "Distribution of Hotel Type",
  x = "Type of Hotel",
  y = "Frequency")
```

```{r resort-t-test, echo = FALSE, message=FALSE}
resort_adr <- hotels[hotels$hotel == "Resort Hotel",]$adr
city_adr <- hotels[hotels$hotel == "City Hotel",]$adr

hotel_type_p_value <- t.test(resort_adr, city_adr)$p.value
print(hotel_type_p_value)
```

We are particularly interested in using the hotel type (Resort versus City Hotel) as a predictor of average daily rate. As an initial analysis, we perform a t-test on the the average daily rates of bookings for resort versus city hotels. This t-test assesses if there is a significant difference between the two distributions, which would indicate that hotel type is likely to be a useful predictor of average daily rate. In this analysis, we find that our t-test produces a p-value of 0.002305%. Because this p-value is very low, we conclude that there is a statistically significant relationship between the average daily hotel rates of resort hotels and city hotels in our dataset, which provides evidence that hotel type is a useful predictor of average daily hotel rates. 

```{r, echo = FALSE, message=FALSE}
hotels %>%
  group_by(hotel) %>%
  summarise(n = n(), mean = mean(adr), sd = sd(adr)) %>%
  kable(digits = 3)
```
We can see that resort hotel is on average more expensive, and that most customers booked at city hotels. We are interested in whether the classification of City Hotel and Resort Hotel will proportionately have similar relationships with the other variables.

```{r adult-t-test,echo = FALSE, message=FALSE}
hotels <-hotels%>%
  mutate(children_present = case_when
         (hotels$adults > 0 & hotels$children == 0 & hotels$babies == 0 ~ 0,
          hotels$children > 0 | hotels$babies > 0 ~ 1), 
         children_present = factor(children_present, 
                          levels = c(0,1)))

adult_adr <- hotels[hotels$children_present == 0,]$adr
kids_adr <- hotels[hotels$children_present == 1,]$adr

resident_type_p_value <- t.test(adult_adr, kids_adr)$p.value
print(resident_type_p_value)
```
We are also particularly interested in using the presence of kids (children or babies) as a predictor of average daily rate. As an initial analysis, we perform a t-test on the the average daily rates of bookings for groups with only adults, versus groups with children or babies. This t-test assesses if there is a significant difference between the two distributions, which would indicate that the presence of kids is likely to be a useful predictor of average daily rate. In this analysis, we find that our t-test produces a p-value of 1.13e-33%. Because this p-value is very low, we conclude that there is a statistically significant relationship between the average daily hotel rates of groups with children or babies and groups with only adults in our dataset, which provides evidence that the presence of kids (children or babies) is a useful predictor of average daily hotel rates. 

# Methodology

## Data Processing

It is important to note that there are a few variables in the dataset that do not provide insight to our research question and as a result we will remove from the dataset. First is the category `hotels$country`, due to the fact that every hotel in this subset of the data is located in the U.S., this is a redundant variable. Next, the variables `hotels$reserved_room_type`, `hotels$assigned_room_type`, `hotels$agent`, `hotels$company` are categorical variables that have codes assigned to their values. However, due to the confidentiality of the customers, the classification of these codes have not been identified by those who collected the data. Therefore, we will be removing these variables from the dataset. In addition to that, we removed `hotels$reservation_status_date` and `hotels$arrival_date_week_number` as we already had variables that tracked day, month, and year. The variables `hotels$previous_cancellations` and `hotels$reservation_status` were also removed as they did not provide relevance in predicting the average daily rate.In addition to removing variables, certain variables such as `hotels$month` as well as `hotels$meal` had to be refactored and cleaned in order to function properly in the linear regression model. 

```{r, echo = FALSE, message=FALSE}

if (hotels$arrival_date_week_number >=36 & hotels$arrival_date_week_number<=39)
{hotels$arrival_date_month = "September"}
if (hotels$arrival_date_week_number >=1 & hotels$arrival_date_week_number<=4)
{hotels$arrival_date_month = "January"}

hotels <- na.omit(hotels)
hotels <- subset(hotels, select = -country)
hotels <- subset(hotels, select = -reserved_room_type)
hotels <- subset(hotels, select = -assigned_room_type)
hotels <- subset(hotels, select = -agent)
hotels <- subset(hotels, select = -company)
hotels <- subset(hotels, select = -reservation_status_date)
hotels <- subset(hotels, select = -previous_cancellations)
hotels <- subset(hotels, select = -reservation_status)
hotels <- subset(hotels, select = -arrival_date_week_number)


  
hotels <- hotels%>%
  mutate(arrival_date_month = factor(arrival_date_month, 
  levels = c("January", "February", "March", "April",
           "May", "June", "July", "August",
           "September", "October", "November",
           "December")), 
        arrival_date_year = factor(arrival_date_year, 
                                     levels = c(2015, 2016, 2017)),
          meal = case_when(meal == "SC" | meal == "Undefined" ~ "None",
                           meal == "BB" ~ "BB",
                           meal == "HB" ~ "HB"),
          meal = factor(meal, levels = c("None", "BB", "HB")))

colnames(hotels)[colnames(hotels) == 'arrival_date_year'] <- 'year'
colnames(hotels)[colnames(hotels) == 'arrival_date_month'] <- 'month'
colnames(hotels)[colnames(hotels) == 'arrival_date_day_of_month'] <- 'day_of_month'
```

We will now fit a model using all of the variables from our cleaned dataset. In particular, we are curious in how 
hotel (Resort Hotel or City Hotel), stays_in_weekend_nights, stays_in_week_nights, the type of meal plan, as well as the time of the reservatuon as predictor variables. As we explore more of the relationships in our dataset, we will be performing selection criterias to find the best model for prediciting the average daily rate of a hotel reservation. We plan to use multiple linear regression with a variety of combinations of interested predictor variables and their interaction. 

```{r, echo = FALSE, message=FALSE}
set.seed(12121)
hotel_split <- initial_split(hotels, prop = 0.8) #80% in the training set
hotel_train <- training(hotel_split) 
hotel_test <- testing(hotel_split)
hotel_test <- hotel_test[!(hotel_test$meal=='Undefined'),]
rownames(hotel_train) <- NULL
rownames(hotel_test) <- NULL
```

## Model Selection

```{r, echo = FALSE, message=FALSE}
full_model <- lm(adr ~ ., data = hotels)
tidy(full_model) %>%
  kable(digits = 3)
#colnames(hotels)
```

The model above is the full model. With such a high number of predictor variables, we believed it would be best to use a backwards model with BIC as our selection criteria. With this, we are able to narrow down our predictor variables to the most significant ones that can be used to draw concrete conclusions as well as predict daily prices for different types of reservations. Additionally, we were specifically interested in how the month of the reservation and the day in the month had an affect on each other. Therefore, we added this interaction term into our model. The final model is displayed in the Results section.

```{r results = "hide", echo = FALSE, message=FALSE}
selected_model <- stats::step(full_model, scope=formula(full_model), direction="backward", k = log(nrow(hotels)))
```
## Model Conditions 

```{r, echo = FALSE ,message=FALSE}
# Residual + QQ Plots
autoplot(selected_model, which = c(1,2))
```
To check that a linear model is applicable to this dataset, we check the following conditions: Linearity, Constant Variance, Normality, and Independence. 

To check for Linearity, we look at the residuals vs fitted plot. We see that the residuals are randomly scattered, which indicates that the linearity condition is met. 

To check for constant variance, we look at the spread of the residuals in the residuals vs fitted plot. We see that the spread of the residuals is approximately equal as the fitted value increases, indicating that the constant variance condition is satisfied.

To check for normality, we look at the QQ-plot for a linear relationship. In the plot, we see a mostly linear relationship, and additionally recognize that our model is robust to deviations because our number of samples is much larger than 30. Therefore, we conclude that the normality condition is met. 

To check for independence, we look at the study design, and find that there is 
no reason to believe that our samples are not independent from the survey, such that 
individual hotel bookings can be assumed to be independent. 

From this analysis, we conclude that the conditions are met for performing linear regression. 

## Model Diagnostics

```{r, echo = FALSE, message=FALSE}
autoplot(selected_model, which = 1, ncol =1, colour = 'hotel')
```
```{r, echo = FALSE, message=FALSE}
autoplot(selected_model, which = 3, ncol =1, colour = 'hotel')
```
```{r, echo = FALSE, message=FALSE}
autoplot(selected_model, which = 4, ncol =1, colour = 'hotel')
```

```{r, echo = FALSE, message=FALSE, }
#Leverage 

lev_threshold <- 2 * (16+1) / nrow(hotels) 
hotel_aug <- augment(selected_model) 

hotel_aug <- hotel_aug%>%
filter(.hat > lev_threshold)# %>% 
#nrow()


# High Magnitude Residual 
hotel_aug <- hotel_aug%>%
filter(.std.resid < -3 | .std.resid > 3) #%>%
#nrow()

# Influential Point 
hotel_aug <- hotel_aug%>%
filter(.cooksd > 0.5) #%>% 
#nrow()
```

We use leverage and Cook's Distance to identify influential observations in our dataset, and use standardized residuals to identify outliers. 

Leverage is a measure of the distance between an observation's values of the predictor variables and the average values of the predictor variables for the entire data set. We define a high leverage point as having a leverage greater than $ 2(p+1) /n  $, where p is the number of predictors and n is the number of observations. We find 448 observations to be high leverage, and consider them to be
potential influential points. 

Standardized residuals can be used to identify potential outliers, as observations that have standardized residuals of large magnitude don't fit the pattern determined by the regression model. We identify potential outliers as observations with standardized residuals with a magnitude greater than or equal to 3. We find 22 observations to be potential outliers.

Cook's distance is a composite measure of an observation's leverage and standardized residual, and is used to identify influential points. An observation is considered a moderately influential point if it’s Cook’s distance is greater than 0.5. We find 1 observation with a Cook's distance greater than 0.5. Therefore, we can conclude that there is 1 moderately influential point. 

After calculating the model diagnostics, we have found that on observation can be classified as an outlier. However, due to the fact that this is a legitimate observation of our dataset, we have chosen to keep this point in our model.


# Results

## Model 

```{r, echo = FALSE, message=FALSE}
#tidy(selected_model) %>%
 # kable()
```

```{r, echo = FALSE, message=FALSE}
model_interact <- lm(adr ~ hotel + is_canceled + lead_time + year + month + day_of_month + stays_in_week_nights + adults + children + meal + distribution_channel + is_repeated_guest + previous_bookings_not_canceled + days_in_waiting_list + total_of_special_requests + month*day_of_month, data = hotels)

tidy(model_interact, conf.int =.95) %>%
  kable(digits = 3)
```


## Interpretation

The model provides a number of useful predictors for determining how the average daily rate or price of a hotel is affected by its customization. The first aspect when choosing a hotel may be the type of hotel one would like to visit. In this model we looked at the difference between staying in a hotel located in a city/urban setting versus staying in a resort style hotel. According the model, the coefficient for a resort hotel is -5.430. This can be interpreted as if a guest decides to stay in a resort hotel versus a city hotel, they are expected to pay on average \$5.43 less per day , holding all else constant. Because the p-value is very close to 0, we can also conclude that this is a significant statistic. Therefore, when deciding on a travel destination, it would be beneficial to consider the difference in price of a city or resort hotel. 

Another important factor when planning a trip or booking a hotel is when you would like to travel or stay. The model covers a variety of predictors that involve the timing of a booking. For example, when considering which month to travel in, there is a possibility that depending on the month you book your stay, the average daily rate may be more expensive or less depending on the month. For example, the model identifies that the arrival date of a stay in August has a coefficient of 77.880. This can be interpreted as if a guest decides to book a stay during the month of August, they would be expected to pay on average \$77.88 more per day, holding all else constant. Additionally, according to the model, if a guest stayed in a hotel with an arrival date in April, May, June, July, August, September, October, or November, they would be expected to pay on average a higher daily rate than staying in the month of January. This is proved by the fact that the p-value for each coefficient of these months is approximately 0, and the 95% confidence interval does not include 0 and has both thresholds being positive values. On the other hand, the months of February, March, and December all have fairly large p-values and have 0 fall in the 95% confidence interval. Therefore, we can say that these months are not statistically different from the month of January, meaning that there will not be a significant difference in the average daily rate whether a guest stays in January or February, March, or December.

Not only is the month of stay important, but the day within a month may also be an important factor in the price of a hotel stay. The model gives a coefficient of -0.562 for the arrival date day of the month, but has a p-value of 0.208. Because this p-value is larger than the $\alpha$ = 0.05 level, we can say that the date of the month is not statistically significant. This is further proved by the 95% confidence interval (-1.435, 0.312), where 0 falls in the interval. Therefore, the arrival date in relation to the day of the month is not a useful predictor for the average daily rate of a hotel stay. 

Additionally, when looking at the interaction between the day of the month and the month, we can see that maybe staying at a certain time within a certain month may have an effect in the cost. The months of March, July, September, and December all had p-values less than $\alpha$ = 0.05 level, so we can conclude that they are statistically significant. This means that the effect of Month of arrival date on the average daily rate is statistically different for July, September, and December. 

Another attribute that may effect the daily price of the hotel is the number of guests and more specifically, the number of adults in the reservation. The model gives a coefficient of 21.456 and a p-value of 0 showing this coefficient is statistically significant. Therefore, we can conclude that for every additional adult added to the reservation, the average daily price of a hotel will increase on average \$21.46, holding all else constant. From this, a pretty good estimate can be made about how much the daily rate will change with the number of adults under the reservation. This can similarly be applied to the number of children staying under a reservation. For every additional child added to the reservation, the average daily price of a hotel will increase on average \$32.02, holding all else constant. We can be confident in this interpretation as the p-value is approximately 0 and 0 does not fall in the 95% confidence interval. Therefore, more children under a reservation will result in an increase of the average daily rate. 

Lastly, we will discuss how a meal plan can possibly change the average daily price of a hotel stay. In the variable meal, the baseline is if a guest were to get no meal plan. The coefficient for Bed and Breakfast has a coefficient of 21.739 and a p-value of approximately 0. Therefore, we can say that if one were to choose the meal plan Bed and Breakfast, they would pay on average \$21.74 more than if they did not buy a meal plan. Similarly for a Half Board meal plan (breakfast and one other meal), the guest is expected to pay on average $50.50 more per day for their hotel stay if they were to get the Half Board plan versus no meal plan. This interpretation is supported by the approximately 0 p-value for this coefficient. 


## Prediction

In addition to discussing how the different predictors affect the average daily rate, one can also use the model to predict how much their hotel stay may cost in relation to these factors. For example, we can predict the average daily rate for a guest who had a reservation at a Resort Hotel that was not canceled, had 90 days elapsed between the entering date of the booking into the PMS and their arrival date, had an arrival date of February 26, 2016, stayed for 3 weekend nights, consisted of 2 adults and 3 children, got a Half Board meal plan, booked directly from the hotel, is not a repeated guest, has not had previous bookings canceled, had 0 days in the wait list and had a total of 4 special requests. 

```{r, echo = FALSE, message=FALSE}
x0 <- tibble(hotel = "Resort Hotel", is_canceled = 0, lead_time = 90, year = "2016", month = "February",
             day_of_month = 26, stays_in_week_nights = 3, adults = 2, children =3, meal = "HB", distribution_channel = "Direct",
             is_repeated_guest = 0, previous_bookings_not_canceled = 0, days_in_waiting_list = 0, total_of_special_requests = 4)

selected_model %>% 
  predict(x0, interval = "confidence", level = .90) %>%
  kable(digits = 3)
```

Based on the above table, we are 95% confident that the average daily rate for this guest will fall somewhere between \$196.49 and \$224.31. Therefore, we can predict that this guest's reservation will have an average daily rate of about $210.40. This model can be applied to a variety of hypothetical guests and reservations to predict how much one's hotel stay would cost. 

```{r, echo = FALSE, message=FALSE}
train_pred <- tibble(predicted = predict(selected_model, hotel_train)) %>%
                          bind_cols(hotel_train)
train_pred %>%
 rmse(truth = adr, estimate = predicted)
test_pred <- tibble(predicted = predict(selected_model, hotel_test)) %>%
                          bind_cols(hotel_test)
test_pred %>%
  rmse(truth = adr, estimate = predicted)
```

To determine the accuracy of the predictive ability of our model, we use the Root Mean Squared Error 
on both our training and test datasets below. We find a very similar error between 
our training and test sets. For the training set, the model had an RMSE of \$33.767 and the testing set had an RMSE of \$34.714. Given that the RMSEs for the training and test set are very close, we can conclude that the model is generalizable based on these values. 

# Discussion and Conclusion

Our model assumes that hotel bookings are independent from each other, which could 
potentially be broken in limited circumstances, such as competing hotel holiday 
promotions.  

It's important to recognize that our model has several limitations. First, we recognize
that the travel industry (including hotel booking) is sensitive to 
tragic circumstances or disasters, which are not accounted for in our model. For example, 
we recognize that our model is not likely to characterize hotel bookings during COVID, given that our 
data is limited to the years 2015 to 2017 (before COVID was in the U.S.). 
Additionally, we anticipate that the hotel room type (such as suite versus 
individual room) would be an important predictor of hotel room price, but is not 
included in our model, as described in our methodologies section.
Additionally, in our exploratory analysis we found that there were far more hotel bookings in 
cities than in resorts. Given this skew, our model may better characterize city 
bookings than resort bookings. 

Our analysis shows that purchase of a meal plan, booking in August or October, and having previous
bookings that were not cancelled were the
most correlated with larger increases in daily hotel rate, and that being a repeated 
guest, booking in a Resort hotel instead of a City hotel, and booking in December were most correlated with decreases in daily hotel rate. We were also able to show that our model generalized to new
data,showing that the error rate on a training and test set are similar (around $33). This
means that our model could reasonably be used to predict hotel rates in the U.S., 
with the above limitations in mind. 


# References


<div id="refs">

</div>


# Appendix

## Exploratory Data Analysis \label{appendix:eda}

```{r, echo = FALSE, message=FALSE}

ggplot(data = hotels, aes(x = meal, fill = meal)) +
geom_bar() + 
labs(title = "Distribution of Meal Plan According to Booking Details",
      y = "Frequency",
      x = "Type of Meal Plan") 
                          
```


```{r, echo = FALSE, message=FALSE}
hotels %>%
  group_by(meal) %>%
  summarise(n = n(), mean = mean(adr), sd = sd(adr)) %>%
  kable(digits = 3)
```

We can see that majority of customer bookings chose the Bed and Breakfast plan (BB), rather than Half Board plan or Full Board (HB and SC). While this does reflect the choice of those booking the hotel, this visualization and dataset can also be influenced by what type of meal plan each hotel is offering. 

```{r, echo = FALSE, message=FALSE}
ggplot(data = hotels, aes(x = lead_time, y = adr)) +
geom_point(color="blue") +
labs(title = "Relationship Between How Early People Book Hotel Rooms
and Average Daily Rate",
  x = "The Number of Days in Advance Hotel Rooms are Booked",
  y = "Average Daily Rate") 
```
We can see that the more last minute visitors book a hotel, the more likely it is that the price in the US dollars varies. Besides, we can also see that the highest price in a hotel decreases as the difference between entering data and arrival date increases. This could suggest that the earlier you book a hotel, the price is more likely to be cheap. 


